### 2020-11-16 main.tex
initial commit

\PassOptionsToPackage{utf8}{inputenc}
%\documentclass{article}

\documentclass[11pt,hidelinks]{article}
%\usepackage[margin=0.75in,top=0.75in,footskip=0.5in]{geometry}
\usepackage{comment}

\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault}

\usepackage[draft]{hyperref}
\usepackage{authblk}
\usepackage{fullpage}

\newcommand{\vocab}{\textbf}

\begin{document}

\title{Hogwild graph genome layout by stochastic gradient descent}

\author{ %[Heumos, \textit{et~al}.]{
  Simon Heumos\,$^{1 \dagger}$,
  Andrea Guarracino\,$^{2 \dagger}$,
  Pjotr Prins\,$^3$,
  and Erik Garrison\,$^{3 *}$
}

\affil{\small$^1$Quantitative Biology Center (QBiC), University of T\"ubingen, T\"ubingen, Germany, 72076 \\
  $^2$University of Rome ``Tor Vergata'', Rome, Italy \\
  $^3$UTHSC, Memphis, TN, USA \\
  $^\ast$To whom correspondence should be addressed. \\
  $^\dagger$Contributed equally.}

\maketitle

\abstract{We sort graphs in 1D and 2D using hogwild stochastic gradient descent (SGD).}

\section{Introduction}

Pangenome graphs \cite{computational2016computational}.

\section{Implementation}

\section{Evaluation}

\section{Discussion}

\section*{Funding}

\bibliography{document}
\bibliographystyle{ieeetr}

\end{document}