### 2022-09-19 main.tex
draft outline

\PassOptionsToPackage{utf8}{inputenc}
%\documentclass{article}

\documentclass[11pt,hidelinks]{article}
%\usepackage[margin=0.75in,top=0.75in,footskip=0.5in]{geometry}
\usepackage{comment}

\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault}

\usepackage[draft]{hyperref}
\usepackage{authblk}
\usepackage{fullpage}

\newcommand{\vocab}{\textbf}

\begin{document}

\title{Pangenome graph layout by HOGWILD! path-guided stochastic gradient descent}

\author{ %[Heumos, \textit{et~al}.]{
  Simon Heumos\,$^{1,2 \dagger}$,
  Andrea Guarracino\,$^{3 \dagger}$,
  Sven Nahnsen\,$^{1,2}$,
  Pjotr Prins\,$^4$,
  and Erik Garrison\,$^{4 *}$
}

\affil{\small$^1$Quantitative Biology Center (QBiC), University of Tübingen, Tübingen 72076, Germany \\
  $^2$ Biomedical Data Science, Department of Computer Science, University of Tübingen , Tübingen 72076, Germany \\ 
  $^3$Genomics Research Centre, Human Technopole, Milan 20157, Italy \\
  $^4$
  Department of Genetics, Genomics and Informatics, University of Tennessee Health Science Center, Memphis, TN 38163, USA \\
  $^\ast$To whom correspondence should be addressed. \\
  $^\dagger$Contributed equally.}

\maketitle

\abstract{High-quality low-cost genome assemblies led to sequencing of whole populations whose genomic variation can be comparatively studied in a graphical pangenome. In pangenome graphs models, DNA sequences are incorporated as nodes with edges connecting the nodes as they occur as sequences representing the graph. Pangenome graphs built from raw sets of alignments may have complex structures generated by common patterns of genome variation. These nonlinear structures can introduce difficulty in downstream analyses, visualization, and interpretation. We propose a new layout algorithm that orders the nodes of a pangenome graph using a HOGWILD! path-guided stochastic gradient descent approach: PG-SGD. Our implementation demonstrates that we can efficiently compute the latent structure of gigabase-scale pangenome graphs revealing their underlying biology.
}
\section{Background}

Pangenome graphs \cite{computational2016computational}. SGD-layout. Layout in general.

\section{Results}

Algorithm 1: How does our node update algorithm work in general, single threaded.
\paragraph{}
Fig. 1: Describe how our approach works. Explanation of 1D graph sorting and metric with Simon's handmade figures. SGD figure. Zipfian distribution figure.
\paragraph{}
Fig. 2: Performance evaluation 1D + 2D: Time + RAM by number of haplotypes. Time + RAM by number of threads. Chr6 HPRC graph?
\\
1D (max. threads) vs. ALIBI; we should compare time + RAM. But also by our sorting goodness metrics: odgi sort + the one suggested in the ALIBI paper. 
\\
2D (max. threads) + odgi draw vs. Bandage.
\paragraph{}
Fig. 3: Cool quantitative 1D sortings and 2D layouts: biological implications. Randomly sorted. PG-SGD sorted. Ygs sorted. We want a pipeline of sortings. 2D layout of the whole HPRC. Chr6 HPRC HLA graph? Chr8 beta-defensin gene cluster HPRC? Whole HPRC?
\paragraph{Bonus Section}
Fig. 4: Detect tension. Relax a graph. Detect tension afterwards. I need to test this on a new data set I got from Erik. I need to establish a fixed lower boundary for the tension from which on we don't relax anymore.

\section{Discussion}

\section{Conclusion}

\section{Methods}

\bibliography{document}
\bibliographystyle{ieeetr}

\section*{Acknowledgements}

\section*{Funding}

\section*{Competing interests}
The authors declare that they have no competing interests.

\section*{Availability of data and materials}

\end{document}