### 2022-09-22 manuscript/fig_algorithm.tex
figure 1 outline

\begin{figure*}[!htb]
	\centering
		%\hline
		\begin{subfigure}[t]{0.4\textwidth}
			\centering
			\caption{}
			\includegraphics[width=\linewidth]{fig/algorithm/TODO_1D_metrics_explained.pdf}
			\label{fig:metrics_explained}
		\end{subfigure}
		\begin{subfigure}[t]{0.4\textwidth}
		\centering
		\caption{}
		\includegraphics[width=\linewidth]{fig/algorithm/TODO_1D_sorting_explained.pdf}
		\label{fig:sorting_explained}
		\end{subfigure}
		%\smallskip
		\begin{subfigure}[t]{0.4\textwidth}
			\centering
			\caption{}
			\includegraphics[width=\linewidth]{fig/algorithm/TODO_pg_sgd_update_sketch.png}
			\label{fig:pg_sgd_update_sketch}
		\end{subfigure}
		\begin{subfigure}[t]{0.4\textwidth}
			\centering
			\caption{}
			\includegraphics[width=\linewidth]{fig/algorithm/TODO_zipf_curve.jpg}
			\label{fig:fig:zipf_curve}
		\end{subfigure}
	\caption{
		Explain metrics and algorithms.
		\textbf{(a)} Sketch of 1D sorting operation. \textbf{(b)} Sketch of metrics calculation. \textbf{(c)} Term updated sketch \textbf{(d)} Zipfian distribution.
	}
	\label{fig:algorithm}
\end{figure*}