### 2023-08-28 manuscript/fig_1d_sorts.tex
add 1D figure, cleaning

\begin{figure*}[!htpb]
	\centering
	\begin{subfigure}{1.0\textwidth}
		\centering
		\caption{}
		\includegraphics[width=1.0\linewidth, trim=0cm 2.5cm 0cm 0cm]{fig/1D/chr6.hprc-v1.0-pggb.13paths.MHC.og.png}
		\label{fig:1d_fig1}
	\end{subfigure}
	\\
	\begin{subfigure}{1.0\textwidth}
		\centering
		\caption{}
		\includegraphics[width=1.0\linewidth, trim=0cm 2.5cm 0cm 0cm]{fig/1D/chr6.hprc-v1.0-pggb.13paths.MHC.og.r.og.png}
		\label{fig:1d_fig2}
	\end{subfigure}
	\\
	\begin{subfigure}{1.0\textwidth}
		\centering
		\caption{}
		\includegraphics[width=1.0\linewidth, trim=0cm 2.5cm 0cm 0cm]{fig/1D/chr6.hprc-v1.0-pggb.13paths.MHC.og.r.og.Y.og.png}
		\label{fig:1d_fig3}
	\end{subfigure}
	\\
	\begin{subfigure}{1.0\textwidth}
		\centering
		\caption{}
		\includegraphics[width=1.0\linewidth, trim=0cm 2.5cm 0cm 0cm]{fig/1D/chr6.hprc-v1.0-pggb.13paths.MHC.og.r.og.HY.og.png}
		\label{fig:1d_fig4}
	\end{subfigure}	
	\caption{\textit{odgi viz} 1D visualizations of a 5 haplotypes subgraph of the Human Pangenome Resource Consortium (HPRC) chromosome 6 pangenome graph. The major histocompatibility complex (MHC) sequence was injected as an extra path. Various node arrangements are shown. 
		\textbf{a-d} A graphs nodes are arranged from left to right forming the pangenome sequence. The black lines under the paths are the links representing the topology. Path names are left.
		\textbf{(a)} \textit{odgi viz} default modality: The colored bars are the paths versus pangenome sequence in a binary matrix. Shown is the subgraph extracted with \textit{odgi extract}. No 1D layout algorithm was applied here.
		\textbf{b-d} \textit{odgi viz} colored by path position. Light grey corresponds to the beginning of a path, black encodes the end of the path.
		\textbf{(b)} The nodes are arranged randomly in 1D.
		\textbf{(c)} The nodes are arranged applying the 1D PG-SGD algorithm.
		\textbf{(d)} The nodes are arranged applying a 1D reference-guided PG-SGD where the nodes of GRCh38 are fixed and only all the othere ones are movable in 1D.}
	\label{fig:1d_sorts}
\end{figure*}